% Ying Wu 2013/06

# Examples of Venn Diagram generation

There are multiple ways to input data for venn diagrams. This document will go over 3 of the common input methods that I have encountered:

## Compressed Vector (Vennerable package)
Each overlap for venn diagram is encoded in a vector where the vector index  
can be converted into a binary number that corresponds to membership in set of venn diagram.


    abc
    000 001 010 011 100 101 110 111
      0   1   2   3   4   5   6   7

`011` represents shared membership / overlap of set b and set c


**Example**
```{r vector, message=FALSE}
library(Vennerable) #https://r-forge.r-project.org/projects/vennerable/
venn_vector = c(0,1,2,3,4,5,6,7) #note the 0 in beginning
plot(Venn(SetNames = c("c", "b", "a"), Weight = venn_vector)) #weighted by default, note set name change
plot(Venn(SetNames = c("c", "b", "a"), Weight = venn_vector), doWeights = FALSE)
# Vennerable can plot up to 5 way venns and has many more parameters, see manual for more details
```

## Binary matrix (limma package)
This is a matrix with columns equal to the number of sets in Venn Diagram 
rows representing each element and TRUE/FALSE representing element membership in each set


    a       b       c
    FALSE   FALSE   TRUE
    FALSE   TRUE    FALSE
    FALSE   TRUE    FALSE
    FALSE   TRUE    TRUE
    FALSE   TRUE    TRUE
    FALSE   TRUE    TRUE
    TRUE    FALSE   FALSE
    TRUE    FALSE   FALSE
    TRUE    FALSE   FALSE
    TRUE    FALSE   FALSE
    ... see appendix for full matrix ...


**Example**
```{r matrix}
library(limma) #http://www.bioconductor.org/packages/release/bioc/html/limma.html
venn_matrix = cbind(a = c(rep(c(FALSE, TRUE), c(6,22))), 
b = c(FALSE, rep(c(TRUE, FALSE, TRUE), c(5,9,13))),
c = c(TRUE, rep(c(FALSE, TRUE, FALSE, TRUE, FALSE, TRUE), c(2:7))))
vennDiagram(vennCounts(venn_matrix))
```

## Identifier list (Venn Diagram package)
This is a list where object in list holds unique IDs that need to be overlapped for venn diagram.


    a
    7 8 9 10 11 12 13 14 15 16 17 18 19 20 21 22 23 24 25 26 27 28
    b
    2 3 4 5 6 16 17 18 19 20 21 22 23 24 25 26 27 28
    c
    1 4 5 6 11 12 13 14 15 22 23 24 25 26 27 28


**Example**
```{r list}
library(VennDiagram)
venn_list = list(a = c(7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28),
b = c(2,3,4,5,6,16,17,18,19,20,21,22,23,24,25,26,27,28),
c = c(1,4,5,6,11,12,13,14,15,22,23,24,25,26,27,28))
grid.draw(venn.diagram(x=venn_list, filename=NULL, scaled=TRUE)) #will draw over current plot
```

## Conversion Function
To convert between the these 3 types I created a function (below) that can be used like this:

    new_vennlist = convertVennInput(venn_matrix, type = "list") #matrix -> list
    new_vennmatrix = convertVennInput(venn_vector, type = "matrix") #vector -> matrix
    new_vennvector = convertVennInput(venn_list, type = "matrix") #list -> vector

Function code below and also [here](convertVennInput.R)

```{r fxn}
convertVennInput = function(venndata, type = "matrix") {
# type can be vector/matrix/list

    type = tolower(type)
    if(type != "vector" & type != "matrix" & type != "list" ) { 
        stop("Invalid type parameter, must be either vector, matrix or list") 
    }

    # convert to matrix-type called venn
    nset = 0
    if(is.list(venndata)) {
        nset = length(venndata)
        tmp = unlist(venndata)
        tmpu = unique(tmp)
        venn = matrix(FALSE, nrow = length(tmpu), ncol = nset)
        colnames(venn) = names(venndata)
        rownames(venn) = sort(tmpu)
        for(i in 1:nset) { venn[,i] = rownames(venn) %in% venndata[[i]] }
    }
    else if(is.vector(venndata)) { #vector
        while(length(venndata) > 2^nset) { nset = nset + 1 } #find number of sets
        if(length(venndata) == 2^nset -1) { 
            venndata = c(0, venndata) #catch case forget 0
        }
        else if(length(venndata) != 2^nset) { 
            stop("Incorrect number of elements in vector") 
        }
        
        venn = matrix(FALSE, nrow = sum(venndata), ncol = nset)
        curRow = venndata[1] + 1
        
        for(i in 2:length(venndata)) {
            venn[curRow:(curRow + venndata[i] -1),c(nset:1)[which(intToBits(i-1) == "01")]] = TRUE
            curRow = curRow + venndata[i]
        }
        colnames(venn) = rev(2^(0:(nset-1)))
    }
    else if(ncol(venndata) > 0) { # matrix (or data frame) might edge cases here
        nset = ncol(venndata)
        if(is.logical(venndata)) { venn = venndata }
        else if(all(venndata == 0 | venndata == 1)) { venn = as.logical(venndata) }
        else { stop("Invalid input data type, must be either vector of numbers, matrix of logicals or list of identifiers") }
    }
    else { stop("Input data type not recognized, must be either vector of numbers, matrix of logicals or list of identifiers") }
    
    #output
    if(type == "matrix") { venn }
    else if(type == "list") { 
        if(is.null(rownames(venn))) { rownames(venn) = 1:nrow(venn) }
        apply(venn, 2, function(i) { rownames(venn)[which(i)] } ) 
    }
    else if(type == "vector") { 
        tmp = venn
        for(i in ncol(venn):1) {
            tmp[venn[,i],i] = 2^(ncol(venn)-i)
        }
        vt = table(rowSums(tmp))
        vec = rep(0, 2^ncol(venn))
        names(vec) = 0:(2^ncol(venn)-1)
        vec[names(vt)] = vt
        vec
    }
}
```

**Notes:**

* Vennerable package can do identifier list -> compressed vector using `Venn()`
* limma package can do binary matrix -> compressed vector using `vennCounts()`

# Appendix
```{r appendix}
venn_vector
venn_matrix
venn_list
sessionInfo()
```

Binary matrix (full)

    a       b       c
    FALSE   FALSE   TRUE
    FALSE   TRUE    FALSE
    FALSE   TRUE    FALSE
    FALSE   TRUE    TRUE
    FALSE   TRUE    TRUE
    FALSE   TRUE    TRUE
    TRUE    FALSE   FALSE
    TRUE    FALSE   FALSE
    TRUE    FALSE   FALSE
    TRUE    FALSE   FALSE
    TRUE    FALSE   TRUE
    TRUE    FALSE   TRUE
    TRUE    FALSE   TRUE
    TRUE    FALSE   TRUE
    TRUE    FALSE   TRUE
    TRUE    TRUE    FALSE
    TRUE    TRUE    FALSE
    TRUE    TRUE    FALSE
    TRUE    TRUE    FALSE
    TRUE    TRUE    FALSE
    TRUE    TRUE    FALSE
    TRUE    TRUE    TRUE
    TRUE    TRUE    TRUE
    TRUE    TRUE    TRUE
    TRUE    TRUE    TRUE
    TRUE    TRUE    TRUE
    TRUE    TRUE    TRUE
    TRUE    TRUE    TRUE


Generated:
```{r timestamp, echo=FALSE, cache=FALSE}
date() #last run
```