% Ying Wu 2013/12
% knit("vennExamples.Rmd"); knit2html("vennExamples.md")
% mv vennExamples.md README.md

# Examples of Venn Diagram generation

There are multiple ways to input data for venn diagrams. This document will go over 3 of the common input methods that I have encountered:

## Compressed Vector (Vennerable package)
Each overlap for venn diagram is encoded in a vector where the vector index  
can be converted into a binary number that corresponds to membership in set of venn diagram.


    abc
    000 001 010 011 100 101 110 111
      0   1   2   3   4   5   6   7

`011` represents shared membership (overlap) of set b and set c


**Example**
```{r vector, message=FALSE}
library(Vennerable) #https://r-forge.r-project.org/projects/vennerable/
venn_vector = c(0,1,2,3,4,5,6,7) #note the 0 in beginning, cannot have name
plot(Venn(SetNames = c("c", "b", "a"), Weight = venn_vector)) #weighted by default, note set name change
plot(Venn(SetNames = c("c", "b", "a"), Weight = venn_vector), doWeights = FALSE)
# Vennerable can plot up to 5 way venns and has many more parameters, see manual for more details
```

## Binary matrix (limma package)
This is a matrix with columns equal to the number of sets in Venn Diagram 
rows representing each element and TRUE/FALSE representing element membership in each set


    a       b       c
    FALSE   FALSE   TRUE
    FALSE   TRUE    FALSE
    FALSE   TRUE    FALSE
    FALSE   TRUE    TRUE
    FALSE   TRUE    TRUE
    FALSE   TRUE    TRUE
    TRUE    FALSE   FALSE
    TRUE    FALSE   FALSE
    TRUE    FALSE   FALSE
    TRUE    FALSE   FALSE
    ... see appendix for full matrix ...


**Example**
```{r matrix}
library(limma) #http://www.bioconductor.org/packages/release/bioc/html/limma.html
venn_matrix = cbind(a = c(rep(c(FALSE, TRUE), c(6,22))), 
b = c(FALSE, rep(c(TRUE, FALSE, TRUE), c(5,9,13))),
c = c(TRUE, rep(c(FALSE, TRUE, FALSE, TRUE, FALSE, TRUE), c(2:7))))
vennDiagram(vennCounts(venn_matrix))
```

## Identifier list (Venn Diagram package)
This is a list where object in list holds unique IDs that need to be overlapped for venn diagram.


    a
    7 8 9 10 11 12 13 14 15 16 17 18 19 20 21 22 23 24 25 26 27 28
    b
    2 3 4 5 6 16 17 18 19 20 21 22 23 24 25 26 27 28
    c
    1 4 5 6 11 12 13 14 15 22 23 24 25 26 27 28


**Example**
```{r list}
library(VennDiagram)
venn_list = list(a = c(7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28),
b = c(2,3,4,5,6,16,17,18,19,20,21,22,23,24,25,26,27,28),
c = c(1,4,5,6,11,12,13,14,15,22,23,24,25,26,27,28))
grid.draw(venn.diagram(x=venn_list, filename=NULL, scaled=TRUE)) #will draw over current plot
```

## Conversion Function
To convert between the these 3 types I created a function [`convertVennInput.R`](convertVennInput.R)

Example:
```{r fxn}
source("convertVennInput.R")

new_vennlist = convertVennInput(venn_matrix, type = "list")     #matrix -> list
new_vennvector = convertVennInput(venn_matrix, type = "vector") #matrix -> vector
new_vennmatrix = convertVennInput(venn_vector, type = "matrix") #vector -> matrix
new_vennlist = convertVennInput(venn_vector, type = "list")     #vector -> list
new_vennvector = convertVennInput(venn_list, type = "vector")   #list -> vector
new_vennmatrix = convertVennInput(venn_list, type = "matrix")   #list -> matrix

# Example of using Venn() from Vennerable package
# Venn() can use identifier list with Set= or
# use compressed vector with weight=
# with vector, need to wrap in as.vector() to remove names
plotVenn(Venn(SetNames = c("a", "b", "c"), Weight = as.vector(new_vennvector)), doWeights = F)
#plot() gives weird error (possibly bug with Vennerable), use plotVenn instead
```

**Notes:**

* Vennerable package can do identifier list -> compressed vector using `Venn(Set=)`
* limma package can do binary matrix -> compressed vector using `vennCounts()`

### Todo
show example of diffbind -> venn diagram

# Appendix
```{r appendix}
venn_vector
venn_matrix
venn_list

new_vennvector
new_vennmatrix
new_vennlist

sessionInfo()
```

Binary matrix (full)

    a       b       c
    FALSE   FALSE   TRUE
    FALSE   TRUE    FALSE
    FALSE   TRUE    FALSE
    FALSE   TRUE    TRUE
    FALSE   TRUE    TRUE
    FALSE   TRUE    TRUE
    TRUE    FALSE   FALSE
    TRUE    FALSE   FALSE
    TRUE    FALSE   FALSE
    TRUE    FALSE   FALSE
    TRUE    FALSE   TRUE
    TRUE    FALSE   TRUE
    TRUE    FALSE   TRUE
    TRUE    FALSE   TRUE
    TRUE    FALSE   TRUE
    TRUE    TRUE    FALSE
    TRUE    TRUE    FALSE
    TRUE    TRUE    FALSE
    TRUE    TRUE    FALSE
    TRUE    TRUE    FALSE
    TRUE    TRUE    FALSE
    TRUE    TRUE    TRUE
    TRUE    TRUE    TRUE
    TRUE    TRUE    TRUE
    TRUE    TRUE    TRUE
    TRUE    TRUE    TRUE
    TRUE    TRUE    TRUE
    TRUE    TRUE    TRUE


Generated:
```{r timestamp, echo=FALSE, cache=FALSE}
date() #last run
```