# Example of how to use makeVenn.R {WIP}

## Intro

I created these function because I got frustrated with how slow the ChIPseqAnno package was and 
the preceieved lack of utilities for making good venn diagrams with GRanges. 

This document is still a work in progress is currently only a brief outline of functions and an example.

## Functions

*peak2GRanges*(bedfile, type="macs", skip=0)
INCOMPLETE. The goal of this function is to convert peak caller output to GRanges
- bedfile is the name file that is peak caller output (typically a bed file)

*createResultMatrix*(typ, fo)
Generate a 'result matrix' (required for every venn diagram)
- results matrix has n columns where n is the number of sets being compared and nrow(fo) rows
- typ is a vector differentiating between the different GRanges
- fo is result of `findOverlaps()`

*extractOverlap*(..., res, typ)
Given res and typ, it will return a true/false vector indexed by whatever column name(s) it is passed in '...'

*printOverlap*(..., res, typ)
Print out the overlaps in a compressed way, need to explain this better when I remember what I did.

*readinGRanges*(...)
Put a bunch of GRanges in, returns result and type matrix `c(res, typ)`

*createOverlapMatrix*(res, typ) 
This function will create an overlap matrix. An overlap matrix is a human readable matrix that enumerates 
all possible overlaps. See source for more details and examples.

*createVenn*(res, typ, overlap = NA, doWeights = FALSE, ...)
Create venn diagram using Vennerable library (can draw up to 5-way venn diagrams, if there are libraries that draw better
I would be happy to stick it in).

*makeVennRunall*(...)
Put in GRanges, get a venn diagram (type and result matrix are returned invisibily)

*makeVennExample*()
INCOMPLETE This example only works if you have named `high1_peaks.bed` `high2_peaks.bed` and `high_peaks.bed` 
in the current directory with score in the 5th column.

## Example

Make a venn diagram from 3 GRanges objects named small/medium/large (in the future `peak2GRanges()` will take over this step)

```{r}
source("makeVenn.R")

# tmpbed = read.table("small.narrowPeak")
    # tmpgrg = GRanges(seqnames = Rle(tmpbed[,1]), ranges = IRanges(start=as.numeric(tmpbed[,2]),
# end=as.numeric(tmpbed[,3]), names=tmpbed[,4]), score=tmpbed[,9])
# small = tmpgrg
# tmpbed = read.table("medium.narrowPeak")
    # tmpgrg = GRanges(seqnames = Rle(tmpbed[,1]), ranges = IRanges(start=as.numeric(tmpbed[,2]),
# end=as.numeric(tmpbed[,3]), names=tmpbed[,4]), score=tmpbed[,9])
# medium = tmpgrg
# tmpbed = read.table("large.narrowPeak")
# tmpgrg = GRanges(seqnames = Rle(tmpbed[,1]), ranges = IRanges(start=as.numeric(tmpbed[,2]),
    # end=as.numeric(tmpbed[,3]), names=tmpbed[,4]), score=tmpbed[,9])
# large = tmpgrg

load("makeVenn.rd")

head(small)
head(medium)
head(large)
```

You can then put it into `makeVennRunall(small, medium, large)` or go through each of the steps individually.
Skip using `readinGRanges()` since this function is not complete and might change later.

```{r}
glg = GRangesList(small, medium, large) #GRanges list
typ = rep(as.character(substitute(list(small, medium, large)))[-1L], as.numeric(lapply(glg, length)))
fo = findOverlaps(unlist(glg), ignoreSelf=T) #find overlaps
res = createResultMatrix(typ, fo) #results matrix
overlap = createOverlapMatrix(res,typ)
createVenn(res, typ, overlap) #will print out overlap matrix
```

Make weighted venn diagram. When the differences in intersections are so great, the venn digram does not look good weighted.
```{r}
createVenn(res, typ, overlap, weighted=TRUE) #will print out overlap matrix
```

Isolate regions in `large` that overlap all regions
```{r}
# large must be specified first in extractOverlap
large_intersect_all = large[extractOverlap("large", "medium", "small", res=res, typ=typ)]
length(large_intersect_all)
head(large_intersect_all)
```

Isolate regions that are unique to `medium`
```{r}
medium_unique = medium[extractOverlap("medium", res=res, typ=typ)]
length(medium_unique)
head(medium_unique)
```

Isolate regions that are self overlaps in `medium`
```{r}
medium_self_overlap = medium[extractOverlap("medium", "medium", res=res, typ=typ)]
length(medium_self_overlap)
head(medium_self_overlap)
```

Note to self: I followed the example shown here: https://github.com/yihui/knitr-examples
